### 2018-04-18 manuscript.Rnw
typose + esteban last fixes

% TODO revert for sumbmisson
\linenumbers
\TODO{revert line numbers before submission}
% TODO also embed bib

@@

The biological significance and determinism of animal behaviours have long been a subject of prime scientific interest.
In the 1960s, behavioural geneticists, such as Seymour Benzer showed that some apparently complex behaviours could be in fact governed by simple genetic determinant, which connected genetics and behaviour\cite{sokolowski_drosophila_2001}.
In the last few decades, our ability to acquire and analyse vast amount of biological data has tremendously increased\cite{stephens_big_2015},
which has been deeply transforming both genetics\cite{schatz_biological_2015} and neuroscience\cite{sejnowski_putting_2014}.
%The rise of high-throughput methods and computational approaches, have particularly transformed genetics and neurobiology.
In fact, ethology itself is undergoing its own transition towards data sciences, which has prompted the terms of `ethomics' \cite{branson_high-throughput_2009, reiser_ethomics_2009} and `computational ethology'\cite{anderson_toward_2014}. 
It is now accepted that the study of behaviour can also benefit from quantitative sciences such as machine learning, physics and computational linguistics \cite{brown_ethology_2018, berman_measuring_2018}.

@@

Although general questions regarding the environmental, evolutionary, neural and genetic determinants of behaviours are shared within the community,
the multiplicity of model organisms, hypotheses and paradigms has led to the existence of a very diverse palette of specific recording techniques.
Some tools were developed, for instance:
to continuously record simple behavioural features such as walking activity\cite{faville_how_2015} and  position\cite{pelkowski_novel_2011} over long durations (days or weeks);
to score more complex ones such as feeding\cite{itskov_automated_2014,ro_flic_2014} and courtship\cite{tsai_image_2012};
and to study the behaviour of multiple interacting animals \cite{swierczek_high-throughput_2011,perez-escudero_idtracker_2014,robie_mapping_2017}.
Whilst most recording platforms are unrelated to each other, there are also some attempts to build general purpose tools that can be adapted by researchers to
suit their specific goals \cite{branson_high-throughput_2009,kabra_jaaba_2013,lopes_bonsai_2015,geissmann_ethoscopes_2017}.
However, when it comes to the subsequent analysis of the generated results, there is still no unified programmatic framework that could be used as a set of building blocks in a pipeline.

The fields of structural biology and  bioinformatics are good examples of communities that have taken advantage of sharing standard files formats, modular command line tools\cite{roumpeka_review_2017} and software packages\cite{huber_orchestrating_2015} that can be assembled into pipelines\cite{leipzig_review_2017}.
In these research areas, which are closely linked to data sciences and statistics, scripting interfaces are the standard since they help to deliver reproducible results \cite{peng_reproducible_2011,stodden_toward_2013}.
In addition, they can be used on remote resources such as computer clusters, which makes them more scalable to the context of `big data'\cite{hashem_rise_2015}.
Since many aspect of behaviour analysis are also becoming increasingly dependent on data sciences, the development of such common tools and data structures would be very valuable.

At first, it may seem as though behavioural experiments are prohibitively heterogeneous -- in terms of model organisms, paradigm and time scale -- for a similar community to arise.
However, some low-level conceptual consistencies and methodological challenges are common across experiments.
For instance, the results (\emph{i.e.} the `data')  feature a set of long time series (sometimes multivariate and irregular), but also contain a formal description of the treatment applied to each individual, the `metadata'.
Storing and accessing data and metadata efficiently involves the implementation of a nested data structure which, in principle,
can be shared between acquisition platforms and experimental paradigms.
% it broaden the scope for developing of utilities an shared principles.

@@

\section*{Design and Implementation}
\texttt{rethomics} is implemented in \texttt{R}\cite{r_core_team_r_2017}, which is widely taught and adopted by computational biologists,
as a collection of packages (Fig~\ref{fig:fig-1}).
Such modular architecture follows the model of modern frameworks such as the \texttt{tidyverse}\cite{wickham_tidyverse_2017}, which results in increased testability, maintainability and adaptability.
In this model, each task of the analysis workflow (\emph{i.e.} data import, manipulation and visualisation) is handled by a different package, and new ones can be designed to suit specific needs.
At the core of \texttt{rethomics} lies the \texttt{behavr} table, a structure used to store large amounts data (\emph{e.g.} position and activity) and metadata (\emph{e.g.} treatment and genotype) in a unique \texttt{data.table}-derived object\cite{dowle_data.table_2017}.
Any input package will import experimental data as a \texttt{behavr} table which can, in turn, be analysed and visualised regardless of the original input platform.
Numerical results and plots are standard objects that can therefore be further analysed inside the wide \texttt{R} package ecosystem.

@@

\begin{figure}[!h]
% 	\includegraphics[width=1\textwidth]{fig/fig-2.pdf}
	\caption{{\bf \texttt{behavr} table.}
	A: Illustration of a \texttt{behavr} object, the core data structure in \texttt{rethomics}.
		The metadata holds a single row for each of the $n$ individuals. 
		Its columns, the $p$ metavariables, are one of two kinds: either required -- and defined by the acquisition platform (\emph{i.e.} used to fetch the data) -- or user-defined (\emph{i.e.} arbitrary).
		In the data, each row is a `read' (\emph{i.e.} information about one individual at one time-point).
		It is formed of $q$ variables and is expected to have a very large number of reads, $k$, for each individual $i$.
		Data and metadata are implicitly joined on the \texttt{id} field.
		Note that the names used for variables and metavariable in this example are only plausible cases which will likely differ in practice. 
		B: Non exhaustive list of uses of a \texttt{behavr} table (referred as \texttt{dt}). 
		In addition to operations on data, which are inherited from \texttt{data.table},
		we provide utilities designed specifically to act on both metadata and data.  
		Commented examples are prefixed by `\texttt{>}'.
	}
	\label{fig:fig-2}
\end{figure}

@@

To integrate visualisation in \texttt{rethomics}, we implemented \texttt{ggetho},
a package that offers new tools that extend the widely adopted \texttt{ggplot2}\cite{wickham_ggplot2_2016}.
\texttt{ggetho} makes full use of the internal \texttt{behavr} structure to summarise temporal trends.
We implemented a set of new `layers' and `scales' that particularly applies to the visualisation of long experiments, with the ability to, for instance, display `double-plotted actograms', periodograms, annotate light and dark phases and wrap time over a given period. 
Importantly, \texttt{ggetho} is fully compatible with \texttt{ggplot2}. 
For instance, \texttt{ggplot2} operations such as faceting, transforming axes and adding new layers will function natively with \texttt{ggetho}.

@@

\section*{Availability and Future Directions}
All packages in the \texttt{rethomics} framework are available under the terms of the GPLv3 license and listed at
\href{https://github.com/rethomics}{https://github.com/rethomics/}.
Extensive installation instructions as well as reproducible demos and tutorials are available at
\href{https://rethomics.github.io/}{https://rethomics.github.io/}.
All packages are continuously integrated and unit tested on several versions of \texttt{R} to minimise the risk of present and future issues.

@@

\section*{Acknowledgements}
We would like to thank people who have directly or indirectly contributed to the this manuscript.
In particular, Han Kim, for his invaluable comments on the early versions of \texttt{rethomics} and his dedicated contribution to the tutorials;
Maite Ogueta, for making the DAMS results data available;
Alice French, Hannah Jones, Diana Bicazan and Florencia Fernandez-Chiappe for their comments as early users;
Marcus Gosh and Tara Kane for their feed back on the manuscript;
Brenna Williams, for her help to support multi-beam DAMS;
Patrick Kr{\"a}tschmer, for his time discussing the conceptual framework.

>>>>

% TODO revert for sumbmisson
%\linenumbers
\TODO{revert line numbers before submission}
% TODO also embed bib

@@

The biological significance and determinism of animal behaviours have long been a subject of prime scientific interest.
In the 1960s, behavioural geneticists, such as Seymour Benzer showed that some apparently complex behaviours could be in fact governed by simple genetic determinant, which connected genetics and behaviour\cite{sokolowski_drosophila_2001}.
In the last few decades, our ability to acquire and analyse vast amounts of biological data has tremendously increased\cite{stephens_big_2015},
which has been deeply transforming both genetics\cite{schatz_biological_2015} and neuroscience\cite{sejnowski_putting_2014}.
%The rise of high-throughput methods and computational approaches, have particularly transformed genetics and neurobiology.
In fact, ethology itself is undergoing its own transition towards data sciences, which has prompted the terms of `ethomics' \cite{branson_high-throughput_2009, reiser_ethomics_2009} and `computational ethology'\cite{anderson_toward_2014}. 
It is now accepted that the study of behaviour can also benefit from quantitative sciences such as machine learning, physics and computational linguistics \cite{brown_ethology_2018, berman_measuring_2018}.

@@

Although general questions regarding the environmental, evolutionary, neural and genetic determinants of behaviours are shared within the community,
the multiplicity of model organisms, hypotheses and paradigms has led to the existence of a very diverse palette of specific recording techniques.
Some tools were developed, for instance:
to continuously record simple behavioural features such as walking activity\cite{faville_how_2015} and  position\cite{pelkowski_novel_2011} over long durations (days or weeks);
to score more complex ones such as feeding\cite{itskov_automated_2014,ro_flic_2014}, aggression\cite{dankert_automated_2009} and courtship\cite{tsai_image_2012};
and to study the behaviour of multiple interacting animals \cite{swierczek_high-throughput_2011,perez-escudero_idtracker_2014,robie_mapping_2017}.
Whilst most recording platforms are unrelated to each other, there are also some attempts to build general purpose tools that can be adapted by researchers to
suit their specific goals \cite{branson_high-throughput_2009,kabra_jaaba_2013,lopes_bonsai_2015,geissmann_ethoscopes_2017}.
However, when it comes to the subsequent analysis of the generated results, there is still no unified programmatic framework that could be used as a set of building blocks in a pipeline.

The fields of structural biology and  bioinformatics are good examples of communities that have taken advantage of sharing standard files formats, modular command line tools\cite{roumpeka_review_2017} and software packages\cite{huber_orchestrating_2015} that can be assembled into pipelines\cite{leipzig_review_2017}.
In these research areas, which are closely linked to data sciences and statistics, scripting interfaces are the standard since they help to deliver reproducible results \cite{peng_reproducible_2011,stodden_toward_2013}.
In addition, they can be used on remote resources such as computer clusters, which makes them more scalable to the context of `big data'\cite{hashem_rise_2015}.
Since many aspects of behaviour analysis are also becoming increasingly dependent on data sciences, the development of such common tools and data structures would be very valuable.

At first, it may seem as though behavioural experiments are prohibitively heterogeneous -- in terms of model organisms, paradigm and timescale -- for a similar community to arise.
However, some low-level conceptual consistencies and methodological challenges are common across experiments.
For instance, the results (\emph{i.e.} the `data')  feature a set of long time series (sometimes multivariate and irregular), but also contain a formal description of the treatment applied to each individual, the `metadata'.
Storing and accessing data and metadata efficiently involves the implementation of a nested data structure which, in principle,
can be shared between acquisition platforms and experimental paradigms.
% it broaden the scope for developing of utilities an shared principles.

@@

\section*{Design and Implementation}
\texttt{rethomics} is implemented in \texttt{R}\cite{r_core_team_r_2017}, which is widely taught and adopted by computational biologists,
as a collection of packages (Fig~\ref{fig:fig-1}).
Such modular architecture follows the model of modern frameworks such as the \texttt{tidyverse}\cite{wickham_tidyverse_2017}, which results in increased testability, maintainability and adaptability.
In this model, each task of the analysis workflow (\emph{i.e.} data import, manipulation and visualisation) is handled by a different package, and new ones can be designed to suit specific needs.
At the core of \texttt{rethomics} lies the \texttt{behavr} table, a structure used to store large amounts of data (\emph{e.g.} position and activity) and metadata (\emph{e.g.} treatment and genotype) in a unique \texttt{data.table}-derived object\cite{dowle_data.table_2017}.
Any input package will import experimental data as a \texttt{behavr} table which can, in turn, be analysed and visualised regardless of the original input platform.
Numerical results and plots are standard objects that can, therefore, be further analysed inside the wide \texttt{R} package ecosystem.

@@

\begin{figure}[!h]
% 	\includegraphics[width=1\textwidth]{fig/fig-2.pdf}
	\caption{{\bf \texttt{behavr} table.}
	A: Illustration of a \texttt{behavr} object, the core data structure in \texttt{rethomics}.
		The metadata holds a single row for each of the $n$ individuals. 
		Its columns, the $p$ metavariables, are one of two kinds: either required -- and defined by the acquisition platform (\emph{i.e.} used to fetch the data) -- or user-defined (\emph{i.e.} arbitrary).
		In the data, each row is a `read' (\emph{i.e.} information about one individual at one time-point).
		It is formed of $q$ variables and is expected to have a very large number of reads, $k$, for each individual $i\in [1,n]$.
		Data and metadata are implicitly joined on the \texttt{id} field.
		Note that the names used for variables and metavariable in this example are only plausible cases which will likely differ in practice. 
		B: Non-exhaustive list of uses of a \texttt{behavr} table (referred as \texttt{dt}). 
		In addition to operations on data, which are inherited from \texttt{data.table},
		we provide utilities designed specifically to act on both metadata and data.  
		Commented examples are prefixed by `\texttt{>}'.
	}
	\label{fig:fig-2}
\end{figure}

@@

To integrate visualisation in \texttt{rethomics}, we implemented \texttt{ggetho},
a package that offers new tools that extend the widely adopted \texttt{ggplot2}\cite{wickham_ggplot2_2016}.
\texttt{ggetho} makes full use of the internal \texttt{behavr} structure to summarise temporal trends.
We implemented a set of new `layers' and `scales' that particularly applies to the visualisation of long experiments, with the ability to, for instance, display `double-plotted actograms', periodograms, annotate light and dark phases and wrap time over a given period. 
Importantly, \texttt{ggetho} is fully compatible with \texttt{ggplot2}. 
For instance, \texttt{ggplot2} operations such as faceting, transforming axes and adding new layers functions natively with \texttt{ggetho}.

@@

\section*{Availability and Future Directions}
All packages in the \texttt{rethomics} framework are available under the terms of the GPLv3 license and listed at
\href{https://github.com/rethomics}{https://github.com/rethomics/}.
Extensive installation instructions, as well as reproducible demos and tutorials, are available at
\href{https://rethomics.github.io/}{https://rethomics.github.io/}.
All packages are continuously integrated and unit tested on several versions of \texttt{R} to minimise the risk of present and future issues.

@@

\section*{Acknowledgements}
We would like to thank people who have directly or indirectly contributed to the this manuscript.
In particular, Han Kim, for his invaluable comments on the early versions of \texttt{rethomics} and his dedicated contribution to the tutorials;
Maite Ogueta, for making the DAMS results data available;
Alice French, Hannah Jones, Diana Bicazan and Florencia Fernandez-Chiappe for their comments as early users;
Marcus Gosh and Tara Kane for their feedback on the manuscript;
Brenna Williams, for her help to support multi-beam DAMS;
Patrick Kr{\"a}tschmer, for his time discussing the conceptual framework.

### 2018-04-18 manuscript.tex
typose + esteban last fixes

In order to illustrate the usefulness of \texttt{rethomics}, we provide two carefully annotated examples.
The first one is a detailed and reproducible description of the loading and analysis activity data in the context of circadian rhythm, using DAM2 (Trikinetics Inc.) data.
The second one shows how \texttt{rethomics} integrates with the rest of \texttt{R} to perform a multi-scale analysis of a periodic behaviour, using continuous wavelet transform, on data generated with ethoscopes\cite{geissmann_ethoscopes_2017}.

\subsection*{Canonical circadian analysis in \emph{Drosophila}}
The \texttt{zeitgebr} package implements a comprehensive suite of methods to analyse circadian rhythms, 
including the computation of autocorrelograms, $\chi{}^2$\cite{sokolove_chi_1978} and Lomb-Scargle\cite{ruf_lomb-scargle_1999} periodograms,
and peak detection.

@@

Here, we guide the reader through the analysis of two of the genotypes employed in that study;
one control group (NKCC\textsuperscript{ox}/+) and one where NKCC\textsuperscript{ox} is overexpressed in clock neurons (TIM/NKCC\textsuperscript{ox}).
In particular, we outline the necessary steps to analyse two repetitions of the same experiment in which a total of 58 animals were recorded for three to four days in LD and then subjected to constant light for six or seven days.
The \texttt{metadata.csv} file as well as all the associated result files can be downloaded at \href{https://zenodo.org/record/1172980}{https://zenodo.org/record/1172980}.

@@

The activity of dead or escaped animals is falsely scored as long series of zeros,
which may be erroneously interpreted as inactivity (see, for instance, individual labelled \texttt{30} and \texttt{18} in Fig~\ref{fig:fig-3}A).
The \texttt{sleepr} package offers a tool to detect and remove such artefactual data.
It proceeds by detecting the first time an animal is immobile for more than 99~\% of the time (the default) for at least \texttt{time\_window} seconds and then discard any subsequent data.

@@

\begin{figure}[!h]
%	\includegraphics[width=1\textwidth]{fig/fig-4.pdf}
	\caption{{\bf Visualisation of the periodicity in activity of eight selected animals.}
		A: Double-plotted actograms showing all activity during experiment.
		Time is defined relative to the transition from LD to LL (at day 0).
		B: $\chi{}^2$ periodograms over the LL part of the experiment matching the animals in A.
		The blue cross represents the highest peak (if present) above the significance threshold (red line).
		Titles on top of each facet refer to the label allocated to each individual.
		See \nameref{S1-Fig} for all 53 animals.
	}
	\label{fig:fig-4}
\end{figure}

@@

\begin{figure}[!h]
%  	\includegraphics[width=1\textwidth]{fig/fig-5.pdf}
	\caption{{\bf Population statistics on circadian phenotype.}
			A: Average periodograms. 
			      The aggregated relative power of the periodogram of all animals.
			      The solid lines and the shaded areas show population means and their 95\% bootstrap confidence interval, respectively.
			B: Frequencies of rhythmic animals.
			      Number of rhythmic animals (\emph{i.e.} with a significant peak) in each genotypes.
			      Dark and clear fillings indicate rhythmic and arhythmic animals, respectively.
      C: Peak periodicity power and average.
			      Values of the peak period for animals with a significant peak (\emph{i.e.} rhythmic).
			      Individual animals are shown by dots whose size represent relative power of the peak period.
			      The error bars are 95\% bootstrap confidence interval on the population mean.
			      }
	\label{fig:fig-5}
\end{figure}

@@

One of the challenges of behaviour analysis is the `nesting' of events happening over different time scales.
In other words, a behavioural variable can be modulated by the circadian rhythm, but also by co-occurring ultradian and infradian rhythms.
For instance, an animal could have rhythmic bursts of locomotor activity recurring at high frequency (\emph{e.g.} 1~min), but also a circadian  (\emph{i.e.} 24~h) regulation of the same variable.
In this example, both rhythms happen at time scales separated by approximately three orders of magnitudes, which makes them difficult to visualise and integrate in the same analysis.
Being able to keep frequency information over multiple scales is however important in some cases.
In particular, when interested in the frequency modulation of a rhythm by another --
that is, if the periodicity of a high frequency rhythm itself can be function of a lower frequency one.
%For instance, the frequency of rhythmic activity bursts could itself be regulated by another rhythm such as the circadian rhythm.

The problem of understanding time series at different scales in not uncommon in fields such as economics\cite{aguiar-conraria_business_2011}, climate sciences\cite{lau_climate_1995} and ecology\cite{cazelles_wavelet_2008} where variables are governed by multiple underlying rhythms (\emph{e.g.} tidal, daily, yearly and multi-yearly).
One approach is to study a variable of interest in the time/period domain using, for instance, continuous wavelet transform (CWT)\cite{grossmann_decomposition_1984}.
In the context of chronobiology, CWT has been suggested as a tool to investigate ultradian rhythms \cite{leise_wavelet_2013}.

@@

\begin{figure}[!h]
	%  	\includegraphics[width=1\textwidth]{fig/fig-5.pdf}
	\caption{{\bf Wavelet analysis of positional data.}
		A: Raw position data for a representative female (top) and male (bottom) \emph{Drosophila} over five days, in black. 
		The thick coloured lines show the average position every two hours.
		The green rectangles in the background shows the two time windows selected for B and C.
		B: Close up of A, showing position over one hour, in the beginning of the L phase of day 1.
		C: Close up of A, showing position over one hour, in the middle of the L phase of day 1.
		D: Continuous wavelet transform spectrogram for the two representative animals.
		E: Average spectrogram across 40 males and 40 females.
		In D and E, the lines on the right shows the marginal power spectra corresponding to the shown spectrograms (average across all time).
		The male data was collected and described in our previous study \cite{geissmann_ethoscopes_2017} (controls in figure~5M-P) and the females data was acquired in parallel, in the same experimental conditions, but not previously published.
	}
	\label{fig:fig-6}
\end{figure}

@@

Interestingly, in females, this low period pace appears to be frequency modulated during the L phase, suggesting a slower walking speed around ZT6~h.
In contrast, males shows only a high frequency rhythm around the phase transitions (L to D and D to L).
Surprisingly, the peak of high frequency rhythm imply a faster pace in males (approximately 60~s) than in females (approximately 120~s) --
indicating that, when active, males walk faster than females.

>>>>

In order to illustrate the usefulness of \texttt{rethomics}, we provide two examples.
The first one is a detailed and reproducible description of the loading and analysis activity data in the context of circadian rhythm, using DAM2 (Trikinetics Inc.) data.
The second one shows how \texttt{rethomics} integrates within \texttt{R} to perform a multi-scale analysis of a periodic behaviour, using continuous wavelet transform, on data generated with ethoscopes\cite{geissmann_ethoscopes_2017}.

\subsection*{Canonical circadian analysis in \emph{Drosophila}}
The \texttt{zeitgebr} package implements a comprehensive suite of methods to analyse circadian rhythms, 
including the computation of autocorrelograms, $\chi{}^2$\cite{sokolove_chi_1978} and Lomb-Scargle\cite{ruf_lomb-scargle_1999} periodograms,
and automatic peak detection.

@@

Here, we guide the reader through the analysis of two of the genotypes employed in that study;
one control group (NKCC\textsuperscript{ox}/+, which is arrhythmic under LL) and one where NKCC\textsuperscript{ox} is overexpressed in clock neurons (TIM/NKCC\textsuperscript{ox}, which is rhythmic under LL).
In particular, we outline the necessary steps to analyse two repetitions of the same experiment in which a total of 58 animals were recorded for three to four days in LD and then subjected to constant light for six or seven days.
The \texttt{metadata.csv} file and all the associated result files can be downloaded at \href{https://zenodo.org/record/1172980}{https://zenodo.org/record/1172980}.

@@

The activity of dead or escaped animals is scored as long series of zeros,
which may be erroneously interpreted as inactivity (see, for instance, individual labelled \texttt{30} and \texttt{18} in Fig~\ref{fig:fig-3}A).
The \texttt{sleepr} package offers a tool to detect and remove such artefactual data.
It proceeds by detecting the first time an animal is immobile for more than 99~\% of the time (the default) for at least \texttt{time\_window} seconds and then discard any subsequent data.

@@

\begin{figure}[!h]
%	\includegraphics[width=1\textwidth]{fig/fig-4.pdf}
	\caption{{\bf Visualisation of the periodicity in the activity of eight selected animals.}
		A: Double-plotted actograms showing all activity during experiment.
		Time is defined relative to the transition from LD to LL (at day 0).
		B: $\chi{}^2$ periodograms over the LL part of the experiment matching the animals in A.
		The blue cross represents the highest peak (if present) above the significance threshold (red line).
		Titles on top of each facet refer to the label allocated to each individual.
		See \nameref{S1-Fig} for all 53 animals.
	}
	\label{fig:fig-4}
\end{figure}

@@

\begin{figure}[!h]
%  	\includegraphics[width=1\textwidth]{fig/fig-5.pdf}
	\caption{{\bf Population statistics on circadian phenotype.}
			A: Average periodograms. 
			      The aggregated relative power of the periodogram of all animals.
			      The solid lines and the shaded areas show population means and their 95\% bootstrap confidence interval, respectively.
			B: Frequencies of rhythmic animals.
			      Number of rhythmic animals (\emph{i.e.} with a significant peak) in each genotypes.
			      Dark and clear fillings indicate rhythmic and arhythmic animals, respectively.
      C: Peak periodicity power and average.
			      Values of the peak period for animals with a significant peak (\emph{i.e.} rhythmic).
			      Individual animals are shown by dots whose size represent the relative power of the peak period.
			      The error bars are 95\% bootstrap confidence interval on the population mean.
			      }
	\label{fig:fig-5}
\end{figure}

@@

One of the challenges of behaviour analysis is the `nesting' of events happening over different timescales.
In other words, a behavioural variable can be modulated by the circadian rhythm, but also by co-occurring ultradian and infradian rhythms.
For instance, an animal could have rhythmic bursts of locomotor activity recurring at high frequency (\emph{e.g.} 1~min), but also a circadian  (\emph{i.e.} 24~h) regulation of the same variable.
In this example, both rhythms happen at timescales separated by approximately three orders of magnitudes, which makes them difficult to visualise and integrate into the same analysis.
Being able to keep frequency information over multiple scales is however important in some cases.
In particular, when interested in the frequency modulation of a rhythm by another --
that is, if the periodicity of a high-frequency rhythm itself can be a function of a lower frequency one.
%For instance, the frequency of rhythmic activity bursts could itself be regulated by another rhythm such as the circadian rhythm.

The problem of understanding time series at different scales is not uncommon in fields such as economics\cite{aguiar-conraria_business_2011}, climate sciences\cite{lau_climate_1995} and ecology\cite{cazelles_wavelet_2008} where variables are governed by multiple underlying rhythms (\emph{e.g.} tidal, daily, yearly and multi-yearly).
One approach is to study a variable of interest in the time/period domain using, for instance, continuous wavelet transform (CWT)\cite{grossmann_decomposition_1984}.
In the context of chronobiology, CWT has been suggested as a tool to investigate ultradian rhythms \cite{leise_wavelet_2013}.

@@

\begin{figure}[!h]
	%  	\includegraphics[width=1\textwidth]{fig/fig-5.pdf}
	\caption{{\bf Wavelet analysis of positional data.}
		A: Raw position data for a representative female (top) and male (bottom) \emph{Drosophila} over five days, in black. 
		The thick coloured lines show the average position every two hours.
		The green rectangles in the background show the two time windows selected for B and C.
		B: Close up of A, showing position over one hour, starting at the beginning of the L phase of day 1.
		C: Close up of A, showing position over one hour, starting at the middle of the L phase of day 1.
		D: Continuous wavelet transform spectrogram for the two representative animals.
		E: Average spectrogram across 40 males and 40 females.
		In D and E, the lines on the right show the marginal power spectra corresponding to the shown spectrograms (average across all time).
		The male data was collected and described in our previous study \cite{geissmann_ethoscopes_2017} (controls in figure~5M-P) and the females data was acquired in parallel, in the same experimental conditions, but not previously published.
	}
	\label{fig:fig-6}
\end{figure}

@@

Interestingly, in females, this low period pace appears to be frequency modulated during the L phase, suggesting a slower walking speed around ZT6~h.
In contrast, males display only a high-frequency rhythm around the phase transitions (L to D and D to L).
Surprisingly, the peak of high-frequency rhythm implies a faster pace in males (approximately 60~s) than in females (approximately 120~s) --
indicating that, when active, males walk faster than females.

### 2018-04-18 results.Rnw
typose + esteban last fixes

\begin{figure}[!h]
%	\includegraphics[width=1\textwidth]{fig/fig-4.pdf}
	\caption{{\bf Visualisation of the periodicity in activity of eight selected animals.}
		A: Double-plotted actograms showing all activity during experiment.
		Time is defined relative to the transition from LD to LL (at day 0).
		B: $\chi{}^2$ periodograms over the LL part of the experiment matching the animals in A.
		The blue cross represents the highest peak (if present) above the significance threshold (red line).
		Titles on top of each facet refer to the label allocated to each individual.
		See \nameref{S1-Fig} for all \Sexpr{nrow(dt[meta=TRUE])} animals.
	}
	\label{fig:fig-4}
\end{figure}

>>>>

\begin{figure}[!h]
%	\includegraphics[width=1\textwidth]{fig/fig-4.pdf}
	\caption{{\bf Visualisation of the periodicity in the activity of eight selected animals.}
		A: Double-plotted actograms showing all activity during experiment.
		Time is defined relative to the transition from LD to LL (at day 0).
		B: $\chi{}^2$ periodograms over the LL part of the experiment matching the animals in A.
		The blue cross represents the highest peak (if present) above the significance threshold (red line).
		Titles on top of each facet refer to the label allocated to each individual.
		See \nameref{S1-Fig} for all \Sexpr{nrow(dt[meta=TRUE])} animals.
	}
	\label{fig:fig-4}
\end{figure}
